TacticToe: Learning the Game of Tactics

### 2018-01-21 tactictoe_jv.tex
moving related works before the conclusion

\begin{itemize}
\item Proof recording at the tactic level is made more precise. Pattern 
matching constructions, opening of submodules are supported. The different 
recording steps (parsing, unfolding and replaying) are regrouped into a single 
\sml function. So, our recording is more complete and fully automatized.
\item Existing tactic orthogonalization on recorded data is made stronger by 
also including the internal ATP \metis in the orthogonalization 
competition. Because we manage a database of theorems for argument 
prediction, we apply here a pruning step to remove easy theorems.
\item Two types of arguments (list of theorems and terms) can be predicted 
independently of the tactic. Combinatorial explosion in the number of tactics 
is prevented through the introduction of abstracted tactics. As desired, their 
instantiations is dependent on the target goal.
\item MCTS (Monte Carlo tree search) now replaces A* in our proof search 
algorithm. MCTS gives a more balanced feedback to guide proof search by 
comparing subtrees instead of leafs. The policy and evaluation are learned 
through supervised learning.
\item The internal ATP \metis used during our proof search is complemented by 
asynchronous calls to \eprover.
\item First evaluation of \tactictoe and \eprover success rates relative to 
the length of the original proofs.
\item User interaction is enhanced by the minimization and prettification 
applied on the generated proof.
\end{itemize}

The system \textsf{ML4PG}~\cite{DBLP:journals/corr/abs-1212-3618,DBLP:journals/mics/HerasK14} 
groups related proofs thanks to its clustering 
algorithms. It allows \coq users to inspire themselves from similar proofs and 
notice 
duplicated proofs. Our predictions comes from a much more detailed description 
of the open goal.
However, we simply create a single label for each tactic call whereas each of 
its
arguments is treated independently in \textsf{ML4PG}. 
Our choice is motivated by the k-NN algorithm already used in
\holyhammer for the selection of theorems.
%, which can be easly adapted to the selection of tactics but would be improved 
%by a 
%version that can learn to predict tactics and their arguments independently.

The system \textsf{ML4PG}~\cite{DBLP:journals/corr/abs-1212-3618,DBLP:journals/mics/HerasK14} 
groups related proofs thanks to its clustering 
algorithms. It allows \coq users to inspire themselves from similar proofs and 
notice 
duplicated proofs. Our predictions comes from a much more detailed description 
of the open goal.
However, we simply create a single label for each tactic call whereas each of 
its
arguments is treated independently in \textsf{ML4PG}. 
Our choice is motivated by the k-NN algorithm already used in
\holyhammer for the selection of theorems.
%, which can be easly adapted to the selection of tactics but would be improved 
%by a 
%version that can learn to predict tactics and their arguments independently.

@@

\begin{definition}(Tactical search tree)\\
A tactical search tree is a sixtuple $(\mathbb{A},\mathbb{G},\mathbb{T},A,G,T)$ 
that respects the following conditions:
\begin{itemize}
\item $\mathbb{A}$ is a set of nodes, $\mathbb{G}$ is a set of goals 
 and $\mathbb{T}$ is a set of tactics.
\item $A$ is a partial function from $\mathbb{G} \times \mathbb{T}$ to 
$\mathbb{A}$.
\item $T$ is a total function from $\mathbb{G}$ to $P(\mathbb{T})$.
\item $G$ is a total function from $\mathbb{A}$ to $P(\mathbb{G})$.
\item $Support(A) = \lbrace (g,t)\ |\ t \in T(g) \rbrace$.
\item if $t \in T(g)$ and $A(g,t) = a$ then $t(g) = G(a)$.
\end{itemize}

@@

Starting from the bottom parent node, we can construct children of a goal of 
this node by 
choosing a goal $g_i^0$ on this node and choosing tactics $t1,t2,\ldots$ to 
apply to this goal. Here the result produces new nodes each containing a list 
of goals to be proven.

@@

\begin{definition}(Solved node and solved goal)
Let $\mathbb{T}$ be a search tree. The set of solved nodes $\mathbb{A}^\infty$ 
and 
the set of solved goals $\mathbb{G}^\infty$ are defined inductively by:

@@

Since there is a bijection between tactics of the tree search 
and the node they 
produced, any function define on tactics of the tree search can be map to a 
function defined on nodes.

@@

\begin{table}[h!]
\centering\ra{1.3}
\small
\begin{tabular}{ll}
\toprule
  & Solved \\
\midrule
   \eprover &  106 \\
   \tactictoe &  139 \\
   \tactictoe + \metis & 182 \\
   \tactictoe + asynchronous \eprover & 182 \\
%x tactictoe + async holyhammer 10 & 183 \\
   \tactictoe + \metis + asynchronous \eprover & 190 \\

@@

@@

Mention the conflicts.
Theorem orthogonalization is not turn on during instantiation of tacitcs except 
for the metis tactic.
Best parameters + comparison with holyhammer.

@@

>>>>

\begin{itemize}
\item Proof recording at the tactic level is made more precise. Pattern 
matching constructions, opening of submodules are supported. The different 
recording steps (parsing, unfolding and replaying) are regrouped into a single 
\sml function. So, our recording is more complete and fully automatized.
%\item Existing tactic orthogonalization on recorded data is made stronger by 
%also including the internal ATP \metis in the orthogonalization 
%competition. Because we manage a database of theorems for argument 
%prediction, we apply here a pruning step to remove easy theorems.
\item Two types of arguments (list of theorems and terms) can be predicted 
independently of the tactic. Combinatorial explosion in the number of tactics 
is prevented through the introduction of abstracted tactics. As desired, their 
instantiations is dependent on the target goal.
\item MCTS (Monte Carlo tree search) now replaces A* in our proof search 
algorithm. MCTS gives a more balanced feedback to guide proof search by 
comparing subtrees instead of leafs. The policy and evaluation are learned 
through supervised learning.

\item First evaluation of \tactictoe and \eprover success rates relative to 
the length of the original proofs.
\item User interaction is enhanced by the minimization and prettification 
applied on the generated proof.
\item The internal ATP \metis used during proof search can be complemented 
by asynchronous calls to \eprover.
\end{itemize}

@@

\begin{definition}(Tactical search tree)\\
A tactical search tree $\mathfrak{T}$ is a sixtuple 
$(\mathbb{A},\mathbb{G},\mathbb{T},A,G,T)$ 
that respects the following conditions:
\begin{itemize}
\item $\mathbb{A}$ is a set of nodes, $\mathbb{G}$ is a set of goals 
 and $\mathbb{T}$ is a set of tactics.
\item $A$ is a partial function from $\mathbb{G} \times \mathbb{T}$ to 
$\mathbb{A}$.
\item $T$ is a total function from $\mathbb{G}$ to $P(\mathbb{T})$.
\item $G$ is a total function from $\mathbb{A}$ to $P(\mathbb{G})$.
\item $Support(A) = \lbrace (g,t)\ |\ t \in T(g) \rbrace$.
\item if $t \in T(g)$ and $A(g,t) = a$ then $t(g) = G(a)$.
\end{itemize}

@@

Starting from the bottom parent node, we can construct children of a goal of 
this node by 
choosing a goal $g_i^0$ on this node and choosing tactics $t_1,t_2,\ldots$ to 
apply to this goal. Here the result produces new nodes each containing a list 
of goals to be proven.

@@

\begin{definition}(Solved node and solved goal)
Let $\mathfrak{T}$ be a search tree. The set of solved nodes 
$\mathbb{A}^\infty$ 
and 
the set of solved goals $\mathbb{G}^\infty$ are defined inductively by:

@@

%Since the support of A gives us a bijection between tactics of the tree search 
%and the node they 
%produced, any function define on tactics of the tree search can be map to a 
%function defined on nodes.

@@

\begin{table}[h!]
\centering\ra{1.3}
\small
\begin{tabular}{ll}
\toprule
  & Solved \\
\midrule
   \eprover &  106 \\
   \tactictoe &  139 \\
   \tactictoe + \metis & 182 \\
   \tactictoe + \eprover & 182 \\
   %x tactictoe + async holyhammer (10 thread) & 183 \\
   \tactictoe + \metis + \eprover & 190 \\

@@

Slight error in this experiments were eprover predictions have been made from 
the preselected 500 theorems instead of the full library. But even with the 
added overhead, we already see 
the benefit of using stronger external prover than \metis.

@@

Mention the conflicts.
Theorem orthogonalization is not turn on during instantiation of tacitcs except 
for the metis tactic.
Best parameters + comparison with holyhammer.
asynchronous eprover is not used  for fairer comparison with eprover.

@@

The system 
\textsf{ML4PG}~\cite{DBLP:journals/corr/abs-1212-3618,DBLP:journals/mics/HerasK14}

groups related proofs thanks to its clustering 
algorithms. It allows \coq users to inspire themselves from similar proofs and 
notice 
duplicated proofs. Our predictions comes from a much more detailed description 
of the open goal.
However, we simply create a single label for each tactic call whereas each of 
its
arguments is treated independently in \textsf{ML4PG}. 
Our choice is motivated by the k-NN algorithm already used in
\holyhammer for the selection of theorems.
%, which can be easly adapted to the selection of tactics but would be improved 
%by a 
%version that can learn to predict tactics and their arguments independently.

The system 
\textsf{ML4PG}~\cite{DBLP:journals/corr/abs-1212-3618,DBLP:journals/mics/HerasK14}

groups related proofs thanks to its clustering 
algorithms. It allows \coq users to inspire themselves from similar proofs and 
notice 
duplicated proofs. Our predictions comes from a much more detailed description 
of the open goal.
However, we simply create a single label for each tactic call whereas each of 
its
arguments is treated independently in \textsf{ML4PG}. 
Our choice is motivated by the k-NN algorithm already used in
\holyhammer for the selection of theorems.
%, which can be easly adapted to the selection of tactics but would be improved 
%by a 
%version that can learn to predict tactics and their arguments independently.
